### 2021-07-18 state-time.Rmd
Updated State/Time Markdown

---
title: "DV-State-Time"
author: "Vy Nguyen"
date: "7/16/2021"
output:
  pdf_document:
    toc: true
    toc_depth: 3
  html_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r echo=FALSE, message= FALSE, fig.asp=1}

library(openintro)
data("birds")
library(ggplot2)
library(patchwork)
library(ggmosaic)
library("viridis")
library(RColorBrewer) # https://www.datanovia.com/en/blog/top-r-color-palettes-to-know-for-great-data-visualization/  -> Purples, BuPu; Map: BrBG
library(knitr)
library(maps)
library(mapdata)
require(plyr)
library(dplyr)
library(forcats)
library(ggwordcloud)

every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}

usa <- map_data('usa')
us.state <- map_data("state")
df.state <- as.data.frame(cbind(state.abb, state.name))
df.state$state.name <- tolower(df.state$state.name)

options(knitr.table.format = "latex")

theme_set(theme_minimal())
color_pal<-c("seagreen", "steelblue", "darkred", "orange")
color_pal_sky<-c("lightgoldenrod","honeydew","skyblue3")
color_pal_TF<-c("darksalmon","cadetblue")
color_pal_TOD<-c("#fed98e","#fff7bc","#d4b9da","#43a2ca")
theme.title<-theme(
    plot.title = element_text(
      hjust = 0, # 0.5 for center
      size = 12,
      color = "#016c59",
      face = "bold"
    ),
    plot.subtitle = element_text(
      hjust = 0, # 0.5 for center
      size = 10,
      color = "gray50",
      face = "italic"
    )
  )

birds_orig<-birds

factor_phase_flt<-c("Parked", "Taxi", "Take-off run", "Climb", "En Route", "Descent", "Approach", "Landing Roll")
birds$phase_of_flt<-factor(birds$phase_of_flt, levels=factor_phase_flt)

factor_struck<-c("0", "1", "2-10", "11-100", "Over 100")
birds$birds_struck<-factor(birds$birds_struck, levels=factor_struck)

factor_sky<-c("No Cloud", "Some Cloud", "Overcast")
birds$sky <- factor(birds$sky, levels=factor_sky)

factor_effect <- c("None","Other","Aborted Take-off","Engine Shut Down","Precautionary Landing")
birds$effect <- factor(birds$effect, levels=factor_effect)

birds$effectYes<- as.factor(birds$effect != "None")

birds$num_engs <- as.factor(birds$num_engs)

#birds seen: remove one wrong entry
  # which(birds$birds_seen==":-10")
birds$birds_seen[9412] <- NA

factor_seen<-c("None","2-10", "11-100")
birds$birds_seen <- factor(birds$birds_seen, levels=factor_seen) 
  # drop unused level ":-10" by assigning levels again
birds$birds_seenYes <- as.factor(!is.na(birds$birds_seen))
#set Na values in birds seen to "None"
birds[is.na(birds$birds_seen),"birds_seen"] <- "None"

#get rows where birds struck is missing
df <- birds[is.na(birds$birds_struck), 
            c("remarks", "birds_seen", "effect", "birds_struck")]
no_strike <- grepl("NOT A STRIKE|NO STRIKE|NOT A STRKE", df$remarks)
# set birds_struck to 0 for "no strike"
birds[is.na(birds$birds_struck),][no_strike,"birds_struck"] <- "0"

# Amount of na for each column
missing <- sapply(birds, function(x) sum(is.na(x)))
missing_df <- data.frame("column" = names(missing), 
                         "count" = missing, row.names = NULL)
names(missing_df)[2] <- "count"
```
\newpage

# The Data

To begin, we first have a look at the dataset we're working with. This is the *birds* dataset form the *openintro* package. It is a collection of all collisions between aircraft in wildlife that were reported to the US Federal Aviation Administration between 1990 and 1997, with details on the circumstances of the collision. It consists of 19302 observations and 17 variables which are given in the table below.

```{r echo=FALSE}
#knitr::opts_chunk$set(tidy.opts=list(width.cutoff=10), tidy=TRUE)
#str(birds_orig)
data_table<-data.frame(Variable = c("opid", "operator", "atype", "remarks", "phase_of_flt", "ac_mass", "num_engs", "date", "time_of_day", "state", "height", "speed", "effect", "sky", "species", "birds_seen", "birds_struck"),
                      Description = c("Three letter identification code for the operator (carrier) of the aircraft.", 
                                      "Name of the aircraft operator.", "Make and model of aircraft.", "Verbal remarks regarding the collision.", "hase of the flight during which the collision occurred.", "Mass of the aircraft classified in kg.", "Number of engines on the aircraft.", "Date of the collision.", "Light conditions", "Two letter abbreviation of the US state in which the collision occurred.", "Feet above ground level.", "Knots (indicated air speed).", "Effect on flight.", "Type of cloud cover, if any.", "Common name for bird or other wildlife.", "Number of birds/wildlife seen by pilot", "Number of birds/wildlife struck."),
                      Type = c("Factor w/ 285 levels", "Factor w/ 285 levels", "Factor w/ 284 levels", "Categorical", "Factor w/ 8 levels", "Discrete (1-5)", "Discrete (1-4)", "Categorical", "Factor w/ 4 levels", "Factor w/ 58 levels", "Continuous (0-32500)", "Continuous (0-400)", "Factor w/ 5 levels", "Factor w/ 3 levels", "Factor w/ 241 levels", "Factor w/ 3 levels", "Factor w/ 5 levels"))
kable(data_table, format="markdown", caption="Variables of the birds dataset")
```

We have 11 factor variables, two categorical ones (with one of them being date information), two discrete numeric variables and two continuous variables.
There are also some missing values which we will look at later.

```{r echo=FALSE, message=FALSE, warning=FALSE}

## Some data exploration

#To start the analysis we will explore the data and look at some of the variables. We can, for example, plot *height* and *speed* as well as the factor variable *birds_struck*:

height.plot<-ggplot(birds) + 
  aes(x = height, y = ..density..) + 
  geom_histogram() +
  geom_density(alpha = 0.25)

speed.plot<-ggplot(birds) + 
  aes(x = speed, y = ..density..) + 
  geom_histogram() +
  geom_density(alpha = 0.25)

struck.plot<-ggplot(birds[!is.na(birds$birds_struck),]) +
  aes(y = birds_struck, x=speed, color=birds_struck, fill=birds_struck) +
  scale_fill_brewer(palette="Blues") +
  scale_color_brewer(palette="Blues") +
  geom_boxplot(alpha=0.2)

time.of.day.plot<-ggplot(birds[!is.na(birds$time_of_day),]) +
  aes(y = time_of_day, x=speed, color=time_of_day, fill=time_of_day) +
  scale_fill_manual(values = color_pal_TOD) +
  scale_color_manual(values = color_pal_TOD) +
  theme(legend.position = "none") +
  geom_boxplot(alpha=0.2)

effect.plot<-ggplot(birds_orig[!is.na(birds$effect),]) +
  aes(x = effect, y=speed, color=effect, fill=effect) +
  scale_fill_brewer(palette="OrRd") +
  scale_color_brewer(palette="OrRd") +
  theme(legend.position = "none") +
  geom_boxplot(alpha=0.2)

dat.plot<-ggplot(birds[!is.na(birds$birds_struck) & birds$birds_struck != "0",]) +
  aes(x = speed, y = height, color=birds_struck, fill=birds_struck) +
  geom_point() +
  facet_wrap(~birds_struck) +
  scale_fill_brewer(palette="Blues") +
  scale_color_brewer(palette="Blues") +
  theme_classic() +
  labs(
    title = "Speed and Height of collisions",
    subtitle = "In relation to the number of wildlife strucked each time",
    x = "Speed in Knots",
    y = "Height in Feet"
  ) +
  theme.title

#(height.plot + speed.plot) / dat.plot
```
\newpage

# Relationship state/wildlife

For the next research question we wanted to explore if there are different kinds of wildlife that aircrafts collide with depending on the state they're in: So essentially the relationship between the State and the Species.

## Species by State

To begin we looked at the **20 most frequent species** in the whole US. The figure shows the number of animals of a specific type which has been displayed in a sorted barplot. We can see that the most frequent species that appear in the data are birds which could not be identified further. Apart from those we see that most bird strikes involve birds with big populations, particularly geese and gulls in the United States. The third most frequent animal, for example, are gulls.

```{r echo=FALSE, fig.asp=0.7}
top20_plot<-ggplot(as.data.frame(sort(table(birds$species), decreasing = TRUE)[1:20]), aes(y=reorder(Var1, Freq), x=Freq)) + 
  geom_bar(stat="identity", fill="darkolivegreen") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) + 
  labs(title="Top 20 in the U.S.",
       subtitle="Species that aircrafts have collided with the most",
       x = "Number of Collisions",
       y = "Species") +
  theme(plot.title.position = "plot") +
  theme.title

# Species in West Virginia collisions
california_collisions<-birds[birds$state == 'WV',]
cal_plot<-ggplot(as.data.frame(sort(table(california_collisions$species), decreasing = TRUE)[1:20]), aes(y=reorder(Var1, Freq), x=Freq)) + 
  geom_bar(stat="identity", fill="darkolivegreen") +
  labs(title="Top 20 in West Virginia",
       subtitle="Most frequent Wildlife species in WV",
       x = "Number of Collisions",
       y = "") +
  theme(plot.title.position = "plot") +
  theme.title

top20_plot + cal_plot
```

When we compare that with the most frequent species in West Virginia (WV), the most frequent species that aircrafts collided with there is surprisingly not a bird but white-tailed deers. And foxes are also among the list. For the rest there are minor differences, for example European Starlings, Rock Pigeons, Sparrows etc. are still one of the most frequent wildlife animals. However, one can see that for example ducks are not that frequent on in WV compared to the nationwide average. Owls also appear more there than in the US in general.

To further examine the state and species variables we could, for example look at all the collisions where gulls where involved and check what states those happened in the most. This is again displayed in barplots:

``` {r echo=FALSE}
# Number of Gulls collisions by state
gulls_collisions <- birds[birds$species == 'GULLS' & !is.na(birds$state),]
gulls_plot<-ggplot(as.data.frame(sort(table(gulls_collisions$state), decreasing = TRUE)), aes(y=reorder(Var1, Freq), x=Freq)) + 
  geom_bar(stat="identity", fill = "darkseagreen") + 
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  theme(legend.position = "none") + 
  labs(title="Gulls collisions by State",
       subtitle="Number of aircraft/gulls collisions by State",
       x = "Number of Collisions",
       y = "U.S State") +
  theme.title +
  scale_y_discrete(breaks=every_nth(5))

top5_gulls<-ggplot(as.data.frame(sort(table(gulls_collisions$state), decreasing = TRUE)[1:5]), aes(y=reorder(Var1, Freq), x=Freq)) + 
  geom_bar(stat="identity", fill="darkseagreen") +
  labs(subtitle="Top 10 states with most number of gulls collisions",
       x = "Number of Collisions",
       y = "") + 
  theme.title

(gulls_plot + top5_gulls)
```

When we look at the top states we see that those are all mainly US states that lie on the coastline, for example California, New York and Florida. This therefore makes sense that collisions with gulls happen there a lot since those birds are often found near where the sea is.

## Collisions per State

We also further examined the number of collisions per state which is displayed below.  
For this, we wanted to display our findings in a map for better visualization. In the following we have displayed the number of collisions, so the variable *birds_struck*, depending on the state. Since this is a factor variable we have unclassed it and mapped the levels to the numeric values below. We then grouped the data by state and summed those up to display in the U.S. map.

```{r echo=FALSE, results='asis' }
map_table<-data.frame(factor = c("1", "2", "3", "4", "5"),
                      level = c("0", "1", "2-10", "11-100", "Over 100"),
                      'mapped value' = c("0", "1", "5", "50", "100"))
kable(map_table, format="markdown", caption="Mapping of the factor variable 'birds_struck'")
```

``` {r echo=FALSE}
birds.with.statenames<-merge(x = birds, y = df.state, by.x = "state", by.y = "state.abb")
filtered.birds<-birds.with.statenames[!is.na(birds.with.statenames$birds_struck),]

filtered.birds$struck_num<-unclass(filtered.birds$birds_struck)
filtered.birds$struck_num <- mapvalues(filtered.birds$struck_num, 
                                     from=c(1,2,3,4,5), 
                                     to=c(0,1,5,50,100))

num.collisions.per.state<-aggregate(filtered.birds$struck_num, by=list(region=filtered.birds$state.name), FUN=sum)
dfs.merged<-merge(us.state, num.collisions.per.state, by = "region")

num.coll.map<-ggplot(data=dfs.merged[with(dfs.merged, order(-group, order)), ], aes(x=long, y=lat, fill=x, group=group)) + 
  geom_polygon(color = "white") + 
  scale_fill_distiller(palette="Blues", trans="reverse") +
  guides(fill=guide_colorbar(title="No. of strucks")) + 
  #theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
  #      axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) + 
  labs(x = "Longitude", y = "Latitude") +
  ggtitle('Number of strucks per State') + 
  theme.title +
  coord_fixed(1.3)

num.coll.map
```

The states with the most collisions seem to be California, Florida, New York and Texas. This is reasonable since those states are the most known ones and there is a lot of air traffic going on: People travel a lot to and from those states so it makes sense that there are the most collisions.

We then also included the variable *ac_mass* to look at:

``` {r echo=FALSE}
# Number of collisions per state
state_plot_df<-birds[birds$birds_struck != "0" & !is.na(birds$birds_struck) & !is.na(birds$state) & !is.na(birds$ac_mass),]
state_plot_df$ac_mass_renamed<-revalue(as.factor(state_plot_df$ac_mass), 
     c("1"="2250 kg or less", "2"="2251-5700 kg", "3"="5701-27000 kg", "4"="27001-272000 kg", "5"="above 272000 kg"))
state_plot<-ggplot(data = state_plot_df) +
  aes(y = fct_rev(fct_infreq(state)), fill = ac_mass_renamed) + 
  guides(fill=guide_legend(title="Mass of the aircraft")) + 
  scale_fill_brewer(palette="Purples") +
  geom_bar() + 
  labs(title="Number of collisions by State",
       subtitle="Divided by the mass of the aircraft",
       x = "Number of Collisions",
       y = "U.S State") +
  theme.title + 
  theme(legend.position="right") +
  scale_y_discrete(breaks=every_nth(5))

state_plot
```

Here, what's interesting to observe is that collisions of aircrafts with a high mass (above 272000 kg) seem to mostly happen in the State of Washington, New York and California. This could be due to the fact that those states are frequently flown to and from, so bigger aircrafts like big passenger planes are used there more than in other states. Another reason could be that there's a big aerospace industry in Washington and the Boeing Everett factory. It's said to be the best state for building boeings.

## Bird collisions vs. non-bird collisions

We further investigated by dividing the data into strucks with birds and strucks with non-birds/other wildlife. To do this we created another variable called *is.bird*: When "bird" appeared in the *remarks* or the *species* variable (along with other words), or when the height where the collision happened was more than 7 Feet above ground level, we categorized those collisions as bird collisions. The remaining ones were then classified as non-bird ones. In the following we see wordclouds of the bird and non-bird species, the size of the words in regard to their respective frequency in the data. (Top 30 for each, the number of the words have been limited due to size and readability.)

``` {r echo=FALSE, include=FALSE}
is.bird<-grepl("BIRD|DUCK|PIGEON|HAWK|GULL|DOVE|PARROT|SWAN|GEESE|GOOSE|TURKEY|NIGHTJAR|GREBE|CARDINAL|MARTIN|BLUE JAY|CRANE|WHIP-POOR-WILL|TOWHEE|WARBLER|CORMORANT|HARRIER|BUNTING|PHEASANT|SANDPIPER|PELICAN|OSPREY|ANHINGA|WREN|MAGPIE|MANNIKIN|FINCH|ORIOLE|SNIPE|WIGEON|STARLING|AVOCET|GROUSE|BOBWHITE|CROW|TERN|VULTURE|ROBIN|OWL|FALCON|KILLDEER|GRACKLE|MYNA|PLOVER|MALLARD|SWALLOW|EGRET|KESTREL|LARK|SPARROW|HERON|EAGLE|CHICKADEE|MERGANSER", birds.with.statenames$species) | grepl("BIRD", birds.with.statenames$remarks) | birds.with.statenames$height >= 7
birds.with.statenames$is.bird<-is.bird

# Species that are bird/non bird
birds.with.statenames.filt<-birds.with.statenames[!is.na(birds.with.statenames$is.bird),]

set.seed(43)
is.bird.species<-as.data.frame(table(birds.with.statenames.filt[birds.with.statenames.filt$is.bird,]$species))
bird_wordcloud<-ggplot(is.bird.species[order(is.bird.species$Freq, decreasing = TRUE),][1:30,], aes(label = Var1, size=Freq)) +
  geom_text_wordcloud_area(area_corr_power = 1) + #, mask = png::readPNG(source="bird.png")) +
  labs(title="Frequent bird species") +
  theme(plot.title = element_text(
      vjust = -30,
      hjust = 0.5, # 0.5 for center
      size = 10,
      color = "darkcyan",
      face = "bold"
    )) +
  scale_size_area(max_size = 5)

set.seed(43)
non.bird.species<-as.data.frame(table(birds.with.statenames.filt[!birds.with.statenames.filt$is.bird,]$species))
non_bird_wordcloud<-ggplot(non.bird.species[order(non.bird.species$Freq, decreasing = TRUE),][1:30,], aes(label = Var1, size=Freq)) +
  geom_text_wordcloud_area(area_corr_power = 1) + #, mask = png::readPNG(source="deer.png")) +
  labs(title="Frequent non-bird species") +
  theme(plot.title = element_text(
      vjust = -30,
      hjust = 0.5, # 0.5 for center
      size = 10,
      color = "darkcyan",
      face = "bold"
    )) +
  scale_size_area(max_size = 5)
```
``` {r echo=FALSE}
bird_wordcloud + non_bird_wordcloud
```

We again can observe what we already saw earlier in the barplot for the most frequent species in general: The most frequent birds are unknown, but also gulls and a loth of other bird types can be found hee again. For the non-bird wildlife we see the white-tailed deer again which is, by far, the most frequent one in this category (visualized by the size of the word), followed by coyotes, mule deers and foxes.

Next, we want to find out if there's a difference in bird and non-bird collisions depending on what state the aircraft is in. To do this, we displayed the number of collisions and used the *is.bird* variable we created to divide the data. This is visualized in the maps below:
The left column shows the bird collisions, thr right one the non-bird ones. The upper maps show the ratio, so the proportion of bird/non-bird collisions and total collisions by state. The two bottom maps show the absolute values, so the number of collisions by state, for bird and non-bird encounters.

``` {r echo=FALSE, fig.asp=0.8}
state.coll.bird.nonbird<-as.data.frame.matrix(table(birds.with.statenames.filt$state.name, birds.with.statenames.filt$is.bird))
colnames(state.coll.bird.nonbird)<-c("nonbird", "bird")
state.coll.bird.nonbird <- cbind(state = rownames(state.coll.bird.nonbird), state.coll.bird.nonbird)
rownames(state.coll.bird.nonbird) <- 1:nrow(state.coll.bird.nonbird)

## Birds/Non-Birds in two maps next to each other
map.bird<-ggplot(state.coll.bird.nonbird, aes(map_id = state)) +
  geom_map(aes(fill = bird), map = us.state) +
  scale_fill_distiller(palette = "PuBu", trans="reverse") +
  guides(fill=guide_colorbar(title="No. of collisions")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  labs(x = "Longitude", y = "Latitude") +
  #ggtitle('Bird strucks per State') + 
  theme.title +
  theme(legend.position="bottom") +
   theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  coord_fixed(1.3)

map.non.bird<-ggplot(state.coll.bird.nonbird, aes(map_id = state)) +
  geom_map(aes(fill = nonbird), map = us.state) +
  scale_fill_distiller(palette = "PuBu", trans="reverse") +
  guides(fill=guide_colorbar(title="No. of collisions")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  labs(x = "Longitude", y = "") +
  #ggtitle('Non-bird strucks per State') + 
  theme.title +
  theme(legend.position="bottom") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank(),
        axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
  coord_fixed(1.3)

is.bird.state<-aggregate(birds.with.statenames.filt$is.bird, by=list(region=birds.with.statenames.filt$state.name), FUN=sum)
num.coll.per.state<-as.data.frame(table(birds.with.statenames.filt$state.name))
is.bird.state$x <- is.bird.state$x / num.coll.per.state$Freq
colnames(is.bird.state)<-c("state", "proportion_birds")

is.not.bird.state<-aggregate(!birds.with.statenames.filt$is.bird, by=list(region=birds.with.statenames.filt$state.name), FUN=sum)
is.not.bird.state$x <- is.not.bird.state$x / num.coll.per.state$Freq
colnames(is.not.bird.state)<-c("state", "proportion_non_birds")

map.bird.prop<-ggplot(is.bird.state, aes(map_id = state)) +
  geom_map(aes(fill = proportion_birds), map = us.state) +
  scale_fill_distiller(palette = "BuGn", trans="reverse") +
  guides(fill=guide_colorbar(title="Ratio")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  labs(x = "", y = "Latitude") +
  ggtitle('Bird collisions per State') + 
  theme.title +
  theme(legend.position="top") +
  coord_fixed(1.3)

map.non.bird.prop<-ggplot(is.not.bird.state, aes(map_id = state)) +
  geom_map(aes(fill = proportion_non_birds), map = us.state) +
  scale_fill_distiller(palette = "BuGn", trans="reverse") +
  guides(fill=guide_colorbar(title="Ratio")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  labs(x = "", y = "") +
  ggtitle('Non-bird collisions per State') + 
  theme.title +
  theme(legend.position="top") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  coord_fixed(1.3)

(map.bird.prop + map.non.bird.prop) / (map.bird + map.non.bird)

########### Ratio of nonbird/bird collisions doesn't match
state.coll.nonbird<-select(state.coll.bird.nonbird, state, nonbird)
state.coll.nonbird$is.bird<-FALSE
colnames(state.coll.nonbird)[2]<-"Freq"
state.coll.bird<-select(state.coll.bird.nonbird, state, bird)
state.coll.bird$is.bird<-TRUE
colnames(state.coll.bird)[2]<-"Freq"
state.coll.bird.nonbird<-rbind(state.coll.nonbird, state.coll.bird)

nbp<-ggplot(state.coll.bird.nonbird, aes(map_id = state)) +
  geom_map(aes(fill = Freq), map = us.state) +
  scale_fill_distiller(palette = "BrBG") +
  guides(fill=guide_colorbar(title="No. of strucks")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  facet_wrap( ~ is.bird)
###########
```

We see that, all in all, the states that have the most bird strucks also have the most non-bird strucks. Though states in the midwest and that do not lie on the coastline have a higher ratio of non-bird collisions than those that do.  
And, what was already quite obvious in the beginning: There are a lot more aircraft-wildlife collisions with birds than other animals (Over 90%). West Virginia has the lowest ratio of all states with approximately 71.28%. This fits with our findings from earlier where we looked at the most frequent species of that state: The white-tailed deer, which is obviously a non-bird animal, was the most common one during aircraft collisions. Though West Virginia doesn't have the most non-bird collisions in absolute values - this spot is taken by Pennsylvania - it has the most in relation to the total number of collisions of the given state.

## Vulture collisions per State

Lastly, for the state/wildlife relationship, we will look at a specific animal, specifically vultures. We look for all collisions with vultures and see which states those happened in. We will look at the proportional values, so the percentage of collisions in a state where vultures were involved in:

```{r echo=FALSE}
is.vulture<-grepl("VULTURE", birds.with.statenames.filt$species)
birds.with.statenames.filt$is.vulture <- is.vulture
is.vulture<-aggregate(birds.with.statenames.filt$is.vulture, by=list(region=birds.with.statenames.filt$state.name), FUN=sum)
is.vulture$x <- is.vulture$x / num.coll.per.state$Freq
colnames(is.vulture)<-c("state", "proportion_vultures")

ggplot(is.vulture, aes(map_id = state)) +
  geom_map(aes(fill = proportion_vultures), map = us.state) +
  scale_fill_distiller(palette = "Reds", trans="reverse") +
  guides(fill=guide_colorbar(title="Ratio of vulture strucks")) + 
  expand_limits(x = us.state$long, y = us.state$lat) +
  labs(x = "Longitude", y = "Latitude") +
  ggtitle('Proportion of vulture strucks per State') + 
  theme.title +
  coord_fixed(1.3)
```

Vultures are more frequent in the South and warm regions which fits with our map above. Their habitat is not in the midwest and hence, there aren't any vulture collisions happening there.

So all in all, we can see that there is a relationship between wildlife and the state. Depending on where the aircraft collided wih wildlife the most frequent species differ from ech other. We can also extract information about the habitat of certain wildlife with the given data (e.g. vulture example). 
\newpage

# Timeline

## The date variable

For the last research question we involved the *date* variable of the dataset. We wanted to investigate the following: Have there been any changes over time regarding the wildlife collisions or are there differences depending on the time of year?  
To do this we first had to perform some feature engineering on *date*: We removed the time part using Regex and formatted it into an actual date variable. We then extracted the month, the year and the weekday into separate variables.

```{r echo=FALSE}
birds$date<-gsub(pattern="\\s?\\d{1,2}:\\d{2}:\\d{2}", replacement="", x=birds$date)
birds$year<-format(as.Date(as.character(birds$date), format = "%m/%d/%Y"), "%Y")
birds$month<-format(as.Date(as.character(birds$date), format = "%m/%d/%Y"), "%m")
birds$day<-format(as.Date(as.character(birds$date), format = "%m/%d/%Y"), "%a")
birds$calendar_week<-format(as.Date(as.character(birds$date), format = "%m/%d/%Y"), "%W")
```
```{r echo=FALSE, results='asis' }
map_date<-data.frame('original date' = c("09/30/1990 0:00:00", "11/29/1993 0:00:00"),
                      date=c("09/30/1990", "11/29/1993"),
                      year = c("1990", "1993"),
                      month = c("09", "11"),
                      day=c("So", "Mo"),
                      calendar_week=c("39", "48"))
kable(map_date, format="markdown", caption="Feature Engineering of the date variable")
```

The above table shows the original *date* variable in the far left column, the other columns are the new variables that *date* has been mapped to.
We now use those to investigate the data further.

## Monthly and weekly grouping

The number of collisions is used as reference against the different date/time variables and also in regards to the *sky*, the *time_of_day* and the *birds_struck* variable.

```{r echo=FALSE}
birds.filt<-birds[!is.na(birds$birds_struck),]

# monthly grouping over all years:
month_plot<-ggplot(birds.filt) +
  aes(x = month, group=birds_struck, fill=birds_struck) +
  scale_fill_brewer(palette="Blues") +
  labs(x = "Month", y = "Number of collisions") +
  #theme(axis.title.y=element_blank()) +
  guides(fill=FALSE) +
  scale_y_continuous(breaks=seq(0,3000,2000)) +
  geom_bar()
# -> more collisions during autumn/fall

# Week of the Year/Calendar Week
woy_plot<-ggplot(birds.filt) +
  aes(x = calendar_week, group=birds_struck, fill=birds_struck) +
  scale_fill_brewer(palette="Blues") +
  labs(x = "Calendar Week", y = "Number of collisions") +
  guides(fill=guide_legend(title="Birds/Wildlife strucked")) +
  geom_bar() +
  scale_y_continuous(breaks=seq(0,600,600)) +
  scale_x_discrete(breaks=every_nth(5))

# Or weekly:
week_plot<-ggplot(birds.filt) + #[as.integer(birds.filt$birds_struck) > 3,]) +
  aes(x = day, group=birds_struck, fill=birds_struck) +
  scale_fill_brewer(palette="Blues") +
  guides(fill=FALSE) +
  labs(x = "Weekday", y = "Prop. number of collisions") +
  geom_bar(position="fill")
# -> bird collisions not dependent on weekdays

month_plot_filt<-ggplot(birds.filt[as.integer(birds.filt$birds_struck) > 3,]) +
  aes(x = month, group=birds_struck, fill=birds_struck) +
  scale_fill_manual(values=c("#3182bd", "#08519c")) +
  guides(fill=FALSE) +
  theme(axis.title.y=element_blank()) +
  labs(x = "Month", y = "Number of collisions") +
  scale_y_continuous(breaks=seq(0, 30, 30)) +
  geom_bar()

birds.filt$first_half_year<-factor(as.factor(as.numeric(birds.filt$month) <= 6), levels=c(TRUE, FALSE))
half.year.plot<-ggplot(birds.filt[as.integer(birds.filt$birds_struck) > 3,]) +
  aes(x = as.factor(first_half_year), group=birds_struck, fill=birds_struck) +
  labs(x = "Is first half of the year", y = "Number of collisions") +
  guides(fill=guide_legend(title="Birds/Wildlife strucked")) +
  theme(legend.position = "none") +
  scale_fill_manual(values=c("#3182bd", "#08519c")) +
  geom_bar()
# way more collisions in second half of the year for birds -> mehr Zugvögel?
# But they would have to come back in spring which we can't see in the diagram?

# Months by sky
birds.filt$month<-format(as.Date(as.character(birds.filt$date), format = "%m/%d/%Y"), "%b")
birds.filt$month<-factor(factor(birds.filt$month), c("Jan","Feb", "Mär", "Apr", "Mai", "Jun", "Jul", "Aug", "Sep", "Okt", "Nov", "Dez"))
xlabels <- c("","02","","04","","06","","08","","10","","12")
#xlabels[seq(1, length(xlabels), 2)] <- ""

sky_month_plot<-ggplot(birds.filt[!is.na(birds.filt$sky),]) +
  aes(x = month, group=sky, fill=sky) +
  scale_fill_manual(values = color_pal_sky) +
  guides(fill=guide_legend(title="Sky")) +
  labs(x = "Month", y = "Prop. number of collisions") +
  geom_bar(position="fill") +
  scale_x_discrete(labels=xlabels)
# -> less overcast during summer months

# Time of Day
tod_plot<-ggplot(birds.filt[!is.na(birds.filt$time_of_day),]) +
  aes(x = month, fill=time_of_day) +
  scale_fill_manual(values = color_pal_TOD) +
  labs(x = "Month", y = "Prop. number of collisions") +
  guides(fill=guide_legend(title="Time of Day")) +
  geom_bar(position="fill") +
  scale_x_discrete(labels=xlabels)
# -> possible explanation: It gets dark later in the summer (in the US)
# -> less strucks during the night in the summer
```
```{r echo=FALSE, fig.asp=0.5}
(sky_month_plot + tod_plot)
```

We can observe that the *sky* variable changes depending on the month: There is less overcast during summer months which is reasonable since generally during summer it rains less.  
We also see that there are less collisions during the night in the summer. One possible explanation here could be that, at least in the U.S., during summer days are longer and it gets darker later. This would also match with the observation we made jut before with the *sky* variable.
So just based on this dataset alone we can also see changes in the weather depending on the season.

```{r echo=FALSE, fig.asp=1.0}
(month_plot / woy_plot) / (week_plot + half.year.plot)

#grid.arrange(month_plot, woy_plot, month_plot_filt, top = "Hawk collisions dover the Year", nrow=3) + theme.title
```

When looking at the number of collisions over the year we can see that there is an increase in collisions during autumn/fall which is especially evident when looking at the collisions where the amount of birds struck was high (where *birds_struck* is "11-100" or "Over 100"). This is displayed in the plot in the bottom right corner: There are way more collisions in the second half of the year than there are in the first half.  
This could be connected to the fact that a lot of birds are migratory birds: So once it gets colder they migrate in big flocks which then leads to more collisions with a higher amount of birds during the fall months. 
However, there is also already an increase during the summer (July, August) and also a slight increase in May, which could again be because of spring migration. For the former, people tend to travel a lot more during summer so this could also explain the increase.  
So with the *birds* dataset we are also able to visualize the migratory behavior of birds over the year.  
We also examined the proportional number of collisions based on the different weekdays but these seem, as expected, to have no influence. So the number of bird/wildlife collisions is not dependent on the days of the week.

## Grouping per Year

For the grouping over the years we will again use the *struck_num* variable from earlier which was based on the *birds_struck* column. We sum up the number of strucked birds/wildlife for each year and display this in a line/point plot. We also do the same again but also divided by each month. Both plots are displayed below.

```{r echo=FALSE, fig.asp=0.8}
birds$struck_num<-unclass(birds$birds_struck)
birds$struck_num <- mapvalues(birds$struck_num, 
                                     from=c(1,2,3,4,5), 
                                     to=c(0,1,5,50,100))

birds.filtered.struck.num<-birds[!is.na(birds$struck_num),]

num.collisions.per.month<-aggregate(birds.filtered.struck.num$struck_num, by=list(month=birds.filtered.struck.num$month), FUN=sum)

strucks.per.year.month<-birds.filtered.struck.num %>%
  group_by(year, month) %>% 
  summarize_at("struck_num",sum,na.rm=TRUE)

coll.year.month<-ggplot(strucks.per.year.month, aes(x = month, y = struck_num, colour=year, group = year)) + 
  labs(x = "Month", y = "Number of strucks", 
       subtitle="(For the years 1990 - 1997)", 
       title = "Number of strucks over the months") +
  theme.title +
  geom_line() + geom_point()

strucks.per.year<-as.data.frame(table(birds.filtered.struck.num$year))
coll.year<-ggplot(strucks.per.year) + 
  aes(x = Var1, y = Freq) + 
  labs(x = "Year", y = "Number of strucks", 
       subtitle="(Aircraft-wildlife collisions from 1990 - 1997)", 
       title = "Number of strucks over the years") +
  theme.title +
  geom_line(group = 1) + geom_point()

coll.year.bar<-ggplot(birds.filtered.struck.num, aes(x = year, group = year)) + geom_bar()

coll.year / coll.year.month #/ coll.year.bar
```

We again see that for almost all years there is an increase in strucked birds in the fall and less collisions in the first half of the year. The reasons for this are the same as above. With each year the number of collisions increases. However, the result for the years "1998" and "1999" are quite low compared to the previous years. After some research we found there was no valid reason for this observation as normally wildlife strikes tend to increase over time as which we have seen with the previous years. Therefore we assume that, at least in this dataset, for the years "1998" and "1999" there are data/reports missing. Especially since the number of collisions for 1999 in total is very close to 0 which is very unlikely to be real.  
Apart from this, our observation fit with general information that can be found with other sources: Wildlife strikes increase with each year and within a year there are more collisions during the fall/bird migration months.

## Hawk collisions over time and per State

Finally, we pick a wildlife species as example: We want to investigate whether we can see changes for hawk collisions depending on the season (summer/winter) and also if there are any significant observations when we also look at the location.  
To do this we look for all entries in the dataset where hawks were involved, so where "hawk" appears in *species*. We then divide the data into two dataframes: One where the month is >= 4 and < 10 (so the summer months from April to September) and the other one with the remaining entries.
For each set we look at the *state* variable too see if the number of strikes changes depending on location. (We again use the *struck_num* variable from earlier.) The values shown are the proportional hawk strikes, so e.g. the hawk strikes in summer in relation to the total hawk strikes per state.

```{r echo=FALSE, fig.asp=1.0}
birds.with.statenames<-merge(x = birds.filtered.struck.num, y = df.state, by.x = "state", by.y = "state.abb")

is.hawk<-grepl("HAWK", birds.with.statenames$species)
filtered.hawk <- birds.with.statenames[is.hawk,]

filtered.hawk.summer<-filtered.hawk[as.numeric(filtered.hawk$month) >= 4 & as.numeric(filtered.hawk$month) < 10,]
filtered.hawk.winter<-filtered.hawk[as.numeric(filtered.hawk$month) <= 3 | as.numeric(filtered.hawk$month) >= 10,]

num.hawk.coll.per.state<-as.data.frame(table(filtered.hawk$state.name))
num.collisions.per.state.summer<-aggregate(filtered.hawk.summer$struck_num, by=list(region=filtered.hawk.summer$state.name), FUN=sum)
num.collisions.per.state.summer<-merge(num.collisions.per.state.summer, num.hawk.coll.per.state, by.x = "region", by.y="Var1")
num.collisions.per.state.summer$x <- num.collisions.per.state.summer$x / num.collisions.per.state.summer$Freq
num.collisions.per.state.winter<-aggregate(filtered.hawk.winter$struck_num, by=list(region=filtered.hawk.winter$state.name), FUN=sum)
num.collisions.per.state.winter<-merge(num.collisions.per.state.winter, num.hawk.coll.per.state, by.x = "region", by.y="Var1")
num.collisions.per.state.winter$x <- num.collisions.per.state.winter$x / num.collisions.per.state.winter$Freq

dfs.merged.summer<-merge(us.state, num.collisions.per.state.summer, by = "region", all.x = TRUE)
dfs.merged.summer$x[is.na(dfs.merged.summer$x)] <- 0
dfs.merged.winter<-merge(us.state, num.collisions.per.state.winter, by = "region", all.x = TRUE)
dfs.merged.winter$x[is.na(dfs.merged.winter$x)] <- 0

plot_summer<-ggplot(data=dfs.merged.summer[with(dfs.merged.summer, order(-group, order)), ], aes(x=long, y=lat, fill=x, group=group)) + 
  geom_polygon(color = "white") + 
  #scale_fill_continuous(low="grey", high="red") +
  scale_fill_distiller(palette = "Reds", trans="reverse") +
  guides(fill=FALSE) + 
  labs(x = "Longitude", y = "Latitude", 
       subtitle="In relation to total hawk trikes", 
       title = "Hawk strikes in Summer") +
  #ggtitle('Hawk collisions during Summer') + 
  theme.title +
  coord_fixed(1.3)

plot_winter<-ggplot(data=dfs.merged.winter[with(dfs.merged.winter, order(-group, order)), ], aes(x=long, y=lat, fill=x, group=group)) + 
  geom_polygon(color = "white") + 
  #scale_fill_continuous(low="grey", high="pink") +
  scale_fill_distiller(palette = "Reds", trans="reverse") +
  guides(fill=guide_colorbar(title="")) +
  labs(x = "Longitude", y = "Latitude", 
       subtitle="In relation to total hawk strikes", 
       title = "Hawk strikes in Winter") +
  theme(axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  #ggtitle('Hawk collisions during Winter') + 
  theme.title +
  coord_fixed(1.3)

#library(gridExtra)
#grid.arrange(plot_summer, plot_winter, top = paste("Hawk collisions depending on the season", "In relation to total hawk collisions", sep="\n"), nrow=1) + theme.title
plot_summer + plot_winter
```

We observe that in the northern U.S. states the proportion of hawk strikes during the summer is the highest. Compared to the winter months the ratio of hawk strikes there is the lowest. And the ratio in the Southeast region increases slightly. And overall there seem to be less hawk strikes during the winter.
This again can be explained with what was mentioned earlier: Hawks are migratory birds as well. During the winter months they travel to the southern regions, hence why there are almost no hawk strikes in the North during winter.

By visualizing our findings in a map and dividing by the date variable we were able to further show the migratory behavior of the birds in the U.S. and how this influences the number of collisions and strikes with aircrafts.